---
title: "Assignment"
author: "Gr"
date: "12/3/2020"
output: html_document
---

##Reading and cleaning initial data
```{r, results = "hide"}
library(lubridate)
unzip(zipfile = "repdata_data_activity.zip")
db <- read.csv("activity.csv", header = TRUE, na.strings = "NA")
db$date <- ymd(db$date)
```

#==================Strategy without fixing NA value===============

## What is a total number of steps during a day avarage across all days for each interval

```{R}

total_steps_day <- aggregate(steps ~ date, data=db, sum)
hist(total_steps_day$steps, breaks = 10, xlab = "number of steps per day", main= NULL)
```
## Mean and median 

```{R}
mean1 <- mean(total_steps_day$steps)
median1 <- median(total_steps_day$steps)
```
Average number of steps per day is `r mean1` and median is `r median1`.


## What is a typical pattern for number os steps for an avarage day and which interval has higest avarege number of steps

```{R}
av_steps_interval <- aggregate(steps ~ interval, data=db, mean)
plot(av_steps_interval)

#define highest number of steps
mx <- max(av_steps_interval$steps)
#select which interval is highest
mx_interval <- av_steps_interval[av_steps_interval$steps == mx, 1]
#adding marker to the plot
abline(v = mx_interval, col = "red")
abline(h = mx, col = "red")

```

#ggplot alternatives not used in this analysis 
#library(ggplot2)
#ggplot(av_steps_interval, aes(interval, steps)) + geom_point(alpha = 1/2)
#qplot(interval, steps, data=av_steps_interval, xlab="interval", ylab="steps")


#==================Strategy with fixing NA value===============
##finding where we have NA value
```{r}
na_in_steps <- sum(is.na(db$steps))
na_in_interval <- sum(is.na(db$interval))
na_in_date <- sum(is.na(db$date))
```

##replacing NA with av value for a given interval
```{R}
db_no_na <- db

for (i in c(1:length(db_no_na$steps))) {
    if (is.na(db_no_na[i,1]) == TRUE) {
      db_no_na[i,1] <- av_steps_interval[i,2]
      } 
}
```


## What is total number of steps taken per av day?
```{R}
total_steps_day_NA_fixed <- aggregate(steps ~ date, data=db_no_na, sum)
par(mfrow = c(1,2))
hist(total_steps_day_NA_fixed$steps, breaks = 10, xlab = "number of steps per day, NA fixed", main= NULL)
hist(total_steps_day$steps, breaks = 10, xlab = "number of steps per day", main= NULL)

#Mean and median number of steps taken each day

mean2 <- mean(total_steps_day_NA_fixed$steps)
median2 <- median(total_steps_day_NA_fixed$steps)
```
Average number of steps per day is `r mean2` and median is `r median2`.


## Analyzing pattern across weekdays/weekends

```{R}
#Adding week day to the data and calculating average pattern across Weekday and Weekdays
weekday <- weekdays(db_no_na$date)
db_no_na_w<- cbind(db_no_na, weekday)
av_steps_interval_na_fixed_weekend <- aggregate(steps ~ interval, data=subset(db_no_na_w, db_no_na_w$weekday == "Saturday" | db_no_na_w$weekday == "Sunday" ), mean)
av_steps_interval_na_fixed_week_day <- aggregate(steps ~ interval, data=subset(db_no_na_w, db_no_na_w$weekday != "Saturday" | db_no_na_w$weekday != "Sunday" ), mean)

#ploting results
par(mfrow = c(1,2))
plot(av_steps_interval_na_fixed_weekend, main = "Weekend")
plot(av_steps_interval_na_fixed_week_day, main = "Week day")
```

There is visible difference in an activity pattern between typical Weekend and Week day. 
